---
title: "Geographic Analysis of Champions"
---

```{r include=FALSE, warning=FALSE}
## Packages
library(tidyverse)
library(dplyr)
library(viridis)
library(ggridges)
library(patchwork)
library(readxl)
library(leaflet)
library(plotly)
library(scales)
library(ggplot2)
library(httr)
library(rvest)

## Data Importing and Cleaning  
years_1 <- c(1900:2012, 2014)
years_2 <- c(2015:2019)
importing_data = function(x){
 
  if(str_detect(x, str_c(years_1, collapse = "|"))) {
  read_csv(x, na = c("NULL", "", "0"), col_types = "cicccciiiicc") 
  } 
  
  else if(str_detect(x, str_c(years_2, collapse = "|"))){
    read_csv(x, na = c("NULL", "", "0"), col_types = "cccicccccccccccccccccciiiiccc")
  }
}

boston_df <- 
  tibble(list.files("data", full.names = TRUE)) %>% 
  setNames("file_name") %>% 
  mutate(data = map(file_name, importing_data)) %>% 
  unnest(data) %>% 
  mutate(year = readr::parse_number(file_name),
        city = coalesce(city, residence),
        display_name = str_replace_all(display_name, "[^a-zA-Z0-9]", " ")) %>% 
  mutate(country_residence = replace(country_residence, country_residence == "AHO", "Netherland Antilles")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "ALB", "Albania")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "ALG", "Algeria")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "AND", "Andorra")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "ARG", "Argentina")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Argenti", "Argentina")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "AUS", "Australia")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Austral", "Australia")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "AUT", "Austria")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "BAH", "Bahamas")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "BAR", "Barbados")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Barbado", "Barbados")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "BDI", "Burundi")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "BLR", "Belarus")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "BEL", "Belgium")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "BER", "Bermuda")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "BRA", "Brazil")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "BRN", "Brunei")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "BUL", "Bulgaria")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "CAN", "Canada")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "CAY", "Cayman")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "CHI", "Chile")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "CHN", "China")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "COL", "Colombia")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Colombi", "Colombia")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "CRC", "Costa Rica")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Costa R", "Costa Rica")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "CRO", "Croatia")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "CYP", "Cyprus")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "CZE", "Czech Republic")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Czech R", "Czech Republic")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "DEN", "Denmark")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "DOM", "Dominican Republic")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Dominic", "Dominican Republic")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "ECU", "Ecuador")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "EGY", "Egypt")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "El Salv", "El Salvador")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "ESA", "El Salvador")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "ESP", "Spain")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "EST", "Estonia")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "ETH", "Ethiopia")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Ethiopi", "Ethiopia")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Faroe I", "Faroe Islands")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "FIN", "Finland")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "FLK", "Falkland Islands")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "FRA", "France")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "GBR", "England")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "GER", "Germany")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "GRE", "Greece")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "GRN", "Greenland")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "GUA", "Guatemala")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Guatema", "Guatemala")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "HKG", "Hong Kong")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Hong Ko", "Hong Kong")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "HON", "Honduras")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Hondura", "Honduras")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "HUN", "Hungary")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "INA", "Indonesia")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Indones", "Indonesia")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "IND", "India")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "IRL", "Ireland")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "ISL", "Iceland")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "ISR", "Israel")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "ITA", "Italy")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "JAM", "Jamaica")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "JPN", "Japan")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "JOR", "Jordan")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "KEN", "Kenya")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "KOR", "Korea")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Korea,", "Korea")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "KSA", "Saudi Arabia")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "KUW", "Kuwait")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "LAT", "Latvia")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "LIE", "Liechtenstein")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Liechte", "Liechtenstein")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Lithuan", "Lithuania")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "LTU", "Lithuania")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "LUX", "Luxembourg")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Luxembo", "Luxembourg")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Macao S", "Macao")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Macedon", "Macedonia")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Malaysi", "Malaysia")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "MAR", "Martinique")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Martini", "Martinique")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "MAS", "Malaysia")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "MEX", "Mexico")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "MGL", "Mongolia")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "MLT", "Malta")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "NCA", "Nicaragua")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "NED", "Netherlands")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Netherl", "Netherlands")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "New Zea", "New Zealand")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "NGR", "Nigeria")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "NOR", "Norway")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "NZL", "New Zealand")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "OMA", "Oman")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "PAK", "Pakistan")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Palesti", "Palestine")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "PAN", "Panama")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "PAR", "Paraguay")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Paragua", "Paraguay")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "PER", "Peru")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "PHI", "Philippines")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Philipp", "Philippines")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "POL", "Poland")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "POR", "Portugal")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Portuga", "Portugal")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Puerto", "Puerto Rico")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "QAT", "Qatar")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "ROU", "Romania")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Saudi A", "Saudi Arabia")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "SIN", "Singapore")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Singapo", "Singapore")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "SLO", "Slovenia")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Slovaki", "Slovakia")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Sloveni", "Slovenia")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "SMR", "San Marino")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "South A", "South Africa")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "SRB", "Serbia")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Sri Lan", "Sri Lanka")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "SUI", "Switzerland")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "SVK", "Slovakia")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "SWE", "Sweden")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Switzer", "Switzerland")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "TCA", "Turks and Caicos")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "THA", "Thailand")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Thailan", "Thailand")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "TPE", "Taipei")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "TRI", "Trinidad")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Trinida", "Trinidad")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "TUR", "Turkey")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "TWN", "Taiwan")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "UAE", "United Arab Emirates")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "UGA", "Uganda")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "UKR", "Ukraine")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "United", "United States")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "URU", "Uruguay")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "USA", "United States")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "VEN", "Venezuela")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Venezue", "Venezuela")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "VGB", "Virgin Islands")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "VIE", "Vietnam")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "ZIM", "Zimbabwe")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "RSA", "South Africa")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "RUS", "Russia")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Russian", "Russia")) %>%
  filter(!is.na(display_name)) %>% 
  select(-file_name, -residence, -first_name, -last_name) %>% 
  rename("country_citizenship" = "contry_citizenship" )

## Continent categorization
boston_df$continent[boston_df$country_residence %in% c(
  "Australia", 
  "New Zealand"
  )] <- "Australia"
boston_df$continent[boston_df$country_residence %in% c(
  "Algeria",
  "Burundi",
  "Egypt",
  "Ethiopia",
  "Kenya",  
  "Morocco", 
  "Nigeria,",  
  "South Africa", 
  "Uganda",
  "Zimbabwe"
  )] <- "Africa"
boston_df$continent[boston_df$country_residence %in% c(
 "China", 
  "Hong Kong", 
  "India",            
  "Indonesia",
  "Israel",  
  "Japan",               
  "Jordan", 
  "Korea",   
  "Malaysia",
  "Mongolia",              
  "Nepal",  
  "Oman", 
  "Pakistan",            
  "Palestine", 
  "Philippines", 
  "Qatar",   
  "Russia", 
  "Saudi Arabia",
  "Singapore",  
  "Sri Lanka",  
  "Taipei",               
  "Taiwan",            
  "Thailand",  
  "Vietnam", 
  "Brunei", 
  "Kuwait",                         
  "Macao",
  "United Arab Emirates"                                  
  )] <- "Asia"
boston_df$continent[boston_df$country_residence %in% c(
  "Canada",  
  "Mexico", 
  "Puerto Rico",
  "United States",
  "Bahamas",                        
  "Barbados",
  "Belize",              
  "Bermuda", 
  "Cayman",                
  "Dominican Republic",  
  "Jamaica", 
  "Martinique",
  "Virgin Islands"                                                           
  )] <- "North America"
boston_df$continent[boston_df$country_residence %in% c(
  "Argentina", 
  "Bolivia",                
  "Brazil", 
  "Chile", 
  "Colombia",           
  "Costa Rica",  
  "Ecuador",  
  "El Salvador", 
  "Guatemala",             
  "Honduras", 
  "Nicaragua",  
  "Panama",             
  "Paraguay", 
  "Peru",  
  "Uruguay",            
  "Venezuela",
  "Aruba",
  "Falkland Island",
  "Trinidad"
  )] <- "South America"
boston_df$continent[boston_df$country_residence %in% c(
  "Albania", 
  "Andorra",            
  "Armenia",
  "Austria",
  "Bahrain",
  "Belarus",
  "Belgium",  
  "Bulgaria", 
  "Croatia", 
  "Cyprus",       
  "Czech Republic",              
  "Denmark",   
  "England", 
  "Finland",               
  "France",              
  "Germany",               
  "Greece", 
  "Greenland", 
  "Hungary",              
  "Iceland",
  "Ireland",                             
  "Italy", 
  "Latvia",        
  "Liechtenstein",            
  "Lithuania", 
  "Luxembourg", 
  "Netherland Antilles",          
  "Netherlands", 
  "Norway",    
  "Poland",             
  "Portugal", 
  "Romania",
  "Serbia",              
  "Slovakia",             
  "Slovenia",
  "Spain",            
  "Sweden",          
  "Switzerland", 
  "Turkey",     
  "Turks and Caicos",
  "Ukraine",
  "Bosnia", 
  "Estonia",                
  "Faroe Islands", 
  "Macedonia",             
  "Malta",     
  "Moldova",               
  "Monaco", 
  "San Marino"  
  )] <- "Europe"

## Data cleaning for geo plots
boston_df2 = boston_df %>% 
  filter(year > 1999) %>% 
  filter(overall == 1) %>% 
  filter(gender == "M") %>% 
  select(year, city, state, overall, everything()) %>% 
  drop_na(city) %>% 
  separate(city, into = c("city", "state", "country"), sep = ",") %>% 
  select(-country, -state) %>% 
  writexl::write_xlsx("interactive_map_men.xlsx")
boston_df3 = boston_df  %>% 
  filter(year > 1999) %>% 
  filter(gender_result == 1) %>% 
  filter(gender == "F") %>% 
  select(year, city, state, overall, everything()) %>% 
  drop_na(city) %>% 
  writexl::write_xlsx("interactive_map_women.xlsx")

map_df = read_excel("data/latitude_longitude_winners.xlsx", sheet = 1) %>% 
  select(year, city, latitude, longitude, age, gender, official_time, display_name) %>% 
  rename(place = city) 
map_df2 = read_excel("data/latitude_longitude_winners.xlsx", sheet = 2) %>% 
  select(year, city, latitude, longitude, age, gender, official_time, display_name) %>% 
  rename(place = city) 

men_open = read_excel("data/geo_winners.xlsx", sheet = 4) %>% 
  janitor::clean_names() %>% 
  filter(!(year == 2013)) %>% 
  rename(place = country) %>% 
  separate(official_time, into = c("data", "official_time"), sep = " (?=[^ ]+$)") %>% 
  select(-data) 
lat_long = read_excel("data/lat_long.xlsx") %>% 
  select(-country) %>% 
  rename(place = name)
men_merge <- merge(men_open,lat_long,by="place") %>% 
  rename(display_name = name) %>% 
  mutate(gender = "M") 
men_merge2 = men_merge %>% 
  mutate(age = NA) %>% 
  filter(year > 1900) %>% 
  filter(year < 2000) 
men_total = rbind(men_merge2, map_df)
women_open = read_excel("data/geo_winners.xlsx", sheet = 3) %>% 
  janitor::clean_names() %>% 
  filter(!(year == 2013)) %>% 
  rename(place = country) %>% 
  separate(official_time, into = c("data", "official_time"), sep = " (?=[^ ]+$)") %>% 
  select(-data) 
women_merge = merge(women_open, lat_long, by="place") %>% 
  rename(display_name = name) %>% 
  mutate(gender = "F") 
  
women_merge2 = women_merge %>% 
  mutate(age = NA) %>% 
  filter(year > 1900) %>% 
  filter(year < 2000) 
woman_total = rbind(women_merge2, map_df2) %>% 
  drop_na(latitude)

women_wheelchair = read_excel("data/geo_winners.xlsx", sheet = "women_wheel_chair") %>% 
    janitor::clean_names() %>% 
    filter(!(year == 2013)) %>% 
    separate(official_time, into = c("data", "official_time"), sep = " (?=[^ ]+$)") %>% 
    select(-data) %>% 
    mutate(gender = "F")
men_wheelchair = read_excel("data/geo_winners.xlsx", sheet = "men_wheel_chair") %>% 
    janitor::clean_names() %>% 
    separate(official_time, into = c("data", "official_time"), sep = " (?=[^ ]+$)") %>% 
    select(-data) %>% 
    mutate(gender = "M")
wheelchair_merge = rbind(women_wheelchair, men_wheelchair) %>% 
  rename(place = country)
wheelchair_total = merge(wheelchair_merge, lat_long, by="place") 
winners_bind = rbind(women_merge, men_merge)
```

<br>

# Male Champions Residence 
```{r warning=FALSE}
map_winners = leaflet(men_total) %>% 
  addProviderTiles(providers$CartoDB.Positron) %>% 
  addMarkers(lat = ~latitude, lng = ~longitude,
  popup = paste("Name:", men_total$display_name, "<br>", "Year:", men_total$year,"<br>",  "Official Time:", men_total$official_time, "<br>", "Age:", men_total$age, "<br>", "Gender:", men_total$gender), clusterOptions = markerClusterOptions())
map_winners
```

The continent with the highest number of male winners from 1900 and beyond is North America, specifically the US, followed by Africa. There are 44 winners from the United States, 20 from Kenya, 10 from Japan. A large portion of the winners came from Eldoret and Meru in Kenya. 

<br>

# Female Champions Residence 
```{r warning=FALSE}
map_winners_women = leaflet(woman_total) %>% 
  addProviderTiles(providers$CartoDB.Positron) %>% 
  addMarkers(lat = ~latitude, lng = ~longitude,
  popup = paste("Name:", woman_total$display_name, "<br>", "Year:", woman_total$year,"<br>",  "Official Time:", woman_total$official_time, "<br>", "Age:", woman_total$age, "<br>", "Gender:", woman_total$gender), clusterOptions = markerClusterOptions())
map_winners_women
```

Africa is the continent with the most winners in the female open division, closely followed by Europe, and North America. Most of the female winners in Africa are from Kenya. This map suggests that a participant's proximity to Boston likely does not play a role in their probability of winning the marathon.

<br>

# Official Times and Country of Residence of Champions by Year in the Open Division
```{r warning=FALSE, message=FALSE}
plot2 = winners_bind %>%
  mutate(text_label = 
           str_c("Name: ", display_name, "\nGender: ", gender)) %>% 
  plot_ly(
    x = ~year, y = ~official_time, text = ~text_label, color = ~place,
    type = "scatter") %>% 
  layout(
    title = "Official Times of Male and Female Winners each Year by Winner's Country",
    xaxis = list(title = 'Year'),
    yaxis = list(title = 'Marathon Time'), 
    legend = list(title=list(text='Residence')))
plot2
```

Over time, the champions of the Boston marathon have been getting faster. While prior to 1950, most of the winners were from the United States and Canada, most of the winners are now from Kenya. Additionally, we can see that the fastest marathon times have been from people from Kenya followed by Ethiopia, and the US.

<br>

# Champions Residence in the Wheelchair Division
```{r warning=FALSE, message = FALSE}
map_wheelchair = leaflet(wheelchair_total) %>% 
  addProviderTiles(providers$CartoDB.Positron) %>% 
  addMarkers(lat = ~latitude, lng = ~longitude,
  popup = paste("Name:", wheelchair_total$name, "<br>", "Year:", wheelchair_total$year,"<br>", "Official Time:", wheelchair_total$official_time, "<br>", "Gender:", wheelchair_total$gender), clusterOptions = markerClusterOptions())
map_wheelchair
```

When assessing the map for the wheelchair division, we can see that 25 winners are from the United States, followed by 19 from Europe, and 10 from Africa. While the male and female open division lack winners from Switzerland, there are 15 winners in the wheelchair division from Switzerland. The fastest time is 1:18 in 2017 by Marcel Hug from Switzerland. 

<br>

# Official Times and Country of Residence of Champions by Year in the Wheelchair Division 
```{r echo=FALSE, message=FALSE, warning=FALSE}
plot3 = wheelchair_total %>%
  mutate(text_label = 
           str_c("Name: ", name, "\nGender: ", gender)) %>% 
  plot_ly(
    x = ~year, y = ~official_time, text = ~text_label, color = ~place,
    type = "scatter") %>% 
  layout(
    title = 'Official Times of Winners each Year in Wheel Chair Division',
    xaxis = list(title = 'Year'),
    yaxis = list(title = 'Marathon Time'), 
    legend = list(title=list(text='Residence')))
plot3
```

While there is a downward trend in the official times for the open division, there is no trend in official times for the wheelchair division. Many of the winners pre-2000 are from the United States while many of the winners post-2000 are from South Africa. The lowest time recorded is from an individual in Switzerland. 

