---
title: "Winning Time Trends"
---

```{r include=FALSE, warning=FALSE}
## Packages
library(tidyverse)
library(dplyr)
library(viridis)
library(ggridges)
library(patchwork)
library(readxl)
library(leaflet)
library(plotly)
library(scales)
library(ggplot2)
library(httr)
library(rvest)

## Data Importing and Cleaning  
years_1 <- c(1900:2012, 2014)
years_2 <- c(2015:2019)
importing_data = function(x){
 
  if(str_detect(x, str_c(years_1, collapse = "|"))) {
  read_csv(x, na = c("NULL", "", "0"), col_types = "cicccciiiicc") 
  } 
  
  else if(str_detect(x, str_c(years_2, collapse = "|"))){
    read_csv(x, na = c("NULL", "", "0"), col_types = "cccicccccccccccccccccciiiiccc")
  }
}

boston_df <- 
  tibble(list.files("data", full.names = TRUE)) %>% 
  setNames("file_name") %>% 
  mutate(data = map(file_name, importing_data)) %>% 
  unnest(data) %>% 
  mutate(year = readr::parse_number(file_name),
        city = coalesce(city, residence),
        display_name = str_replace_all(display_name, "[^a-zA-Z0-9]", " ")) %>% 
  mutate(country_residence = replace(country_residence, country_residence == "AHO", "Netherland Antilles")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "ALB", "Albania")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "ALG", "Algeria")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "AND", "Andorra")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "ARG", "Argentina")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Argenti", "Argentina")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "AUS", "Australia")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Austral", "Australia")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "AUT", "Austria")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "BAH", "Bahamas")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "BAR", "Barbados")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Barbado", "Barbados")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "BDI", "Burundi")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "BLR", "Belarus")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "BEL", "Belgium")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "BER", "Bermuda")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "BRA", "Brazil")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "BRN", "Brunei")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "BUL", "Bulgaria")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "CAN", "Canada")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "CAY", "Cayman")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "CHI", "Chile")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "CHN", "China")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "COL", "Colombia")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Colombi", "Colombia")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "CRC", "Costa Rica")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Costa R", "Costa Rica")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "CRO", "Croatia")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "CYP", "Cyprus")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "CZE", "Czech Republic")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Czech R", "Czech Republic")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "DEN", "Denmark")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "DOM", "Dominican Republic")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Dominic", "Dominican Republic")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "ECU", "Ecuador")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "EGY", "Egypt")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "El Salv", "El Salvador")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "ESA", "El Salvador")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "ESP", "Spain")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "EST", "Estonia")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "ETH", "Ethiopia")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Ethiopi", "Ethiopia")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Faroe I", "Faroe Islands")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "FIN", "Finland")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "FLK", "Falkland Islands")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "FRA", "France")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "GBR", "England")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "GER", "Germany")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "GRE", "Greece")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "GRN", "Greenland")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "GUA", "Guatemala")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Guatema", "Guatemala")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "HKG", "Hong Kong")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Hong Ko", "Hong Kong")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "HON", "Honduras")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Hondura", "Honduras")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "HUN", "Hungary")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "INA", "Indonesia")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Indones", "Indonesia")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "IND", "India")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "IRL", "Ireland")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "ISL", "Iceland")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "ISR", "Israel")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "ITA", "Italy")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "JAM", "Jamaica")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "JPN", "Japan")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "JOR", "Jordan")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "KEN", "Kenya")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "KOR", "Korea")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Korea,", "Korea")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "KSA", "Saudi Arabia")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "KUW", "Kuwait")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "LAT", "Latvia")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "LIE", "Liechtenstein")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Liechte", "Liechtenstein")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Lithuan", "Lithuania")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "LTU", "Lithuania")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "LUX", "Luxembourg")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Luxembo", "Luxembourg")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Macao S", "Macao")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Macedon", "Macedonia")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Malaysi", "Malaysia")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "MAR", "Martinique")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Martini", "Martinique")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "MAS", "Malaysia")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "MEX", "Mexico")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "MGL", "Mongolia")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "MLT", "Malta")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "NCA", "Nicaragua")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "NED", "Netherlands")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Netherl", "Netherlands")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "New Zea", "New Zealand")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "NGR", "Nigeria")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "NOR", "Norway")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "NZL", "New Zealand")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "OMA", "Oman")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "PAK", "Pakistan")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Palesti", "Palestine")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "PAN", "Panama")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "PAR", "Paraguay")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Paragua", "Paraguay")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "PER", "Peru")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "PHI", "Philippines")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Philipp", "Philippines")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "POL", "Poland")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "POR", "Portugal")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Portuga", "Portugal")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Puerto", "Puerto Rico")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "QAT", "Qatar")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "ROU", "Romania")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Saudi A", "Saudi Arabia")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "SIN", "Singapore")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Singapo", "Singapore")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "SLO", "Slovenia")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Slovaki", "Slovakia")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Sloveni", "Slovenia")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "SMR", "San Marino")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "South A", "South Africa")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "SRB", "Serbia")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Sri Lan", "Sri Lanka")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "SUI", "Switzerland")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "SVK", "Slovakia")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "SWE", "Sweden")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Switzer", "Switzerland")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "TCA", "Turks and Caicos")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "THA", "Thailand")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Thailan", "Thailand")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "TPE", "Taipei")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "TRI", "Trinidad")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Trinida", "Trinidad")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "TUR", "Turkey")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "TWN", "Taiwan")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "UAE", "United Arab Emirates")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "UGA", "Uganda")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "UKR", "Ukraine")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "United", "United States")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "URU", "Uruguay")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "USA", "United States")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "VEN", "Venezuela")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Venezue", "Venezuela")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "VGB", "Virgin Islands")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "VIE", "Vietnam")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "ZIM", "Zimbabwe")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "RSA", "South Africa")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "RUS", "Russia")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Russian", "Russia")) %>%
  filter(!is.na(display_name)) %>% 
  select(-file_name, -residence, -first_name, -last_name) %>% 
  rename("country_citizenship" = "contry_citizenship" )

## Continent categorization
boston_df$continent[boston_df$country_residence %in% c(
  "Australia", 
  "New Zealand"
  )] <- "Australia"
boston_df$continent[boston_df$country_residence %in% c(
  "Algeria",
  "Burundi",
  "Egypt",
  "Ethiopia",
  "Kenya",  
  "Morocco", 
  "Nigeria,",  
  "South Africa", 
  "Uganda",
  "Zimbabwe"
  )] <- "Africa"
boston_df$continent[boston_df$country_residence %in% c(
 "China", 
  "Hong Kong", 
  "India",            
  "Indonesia",
  "Israel",  
  "Japan",               
  "Jordan", 
  "Korea",   
  "Malaysia",
  "Mongolia",              
  "Nepal",  
  "Oman", 
  "Pakistan",            
  "Palestine", 
  "Philippines", 
  "Qatar",   
  "Russia", 
  "Saudi Arabia",
  "Singapore",  
  "Sri Lanka",  
  "Taipei",               
  "Taiwan",            
  "Thailand",  
  "Vietnam", 
  "Brunei", 
  "Kuwait",                         
  "Macao",
  "United Arab Emirates"                                  
  )] <- "Asia"
boston_df$continent[boston_df$country_residence %in% c(
  "Canada",  
  "Mexico", 
  "Puerto Rico",
  "United States",
  "Bahamas",                        
  "Barbados",
  "Belize",              
  "Bermuda", 
  "Cayman",                
  "Dominican Republic",  
  "Jamaica", 
  "Martinique",
  "Virgin Islands"                                                           
  )] <- "North America"
boston_df$continent[boston_df$country_residence %in% c(
  "Argentina", 
  "Bolivia",                
  "Brazil", 
  "Chile", 
  "Colombia",           
  "Costa Rica",  
  "Ecuador",  
  "El Salvador", 
  "Guatemala",             
  "Honduras", 
  "Nicaragua",  
  "Panama",             
  "Paraguay", 
  "Peru",  
  "Uruguay",            
  "Venezuela",
  "Aruba",
  "Falkland Island",
  "Trinidad"
  )] <- "South America"
boston_df$continent[boston_df$country_residence %in% c(
  "Albania", 
  "Andorra",            
  "Armenia",
  "Austria",
  "Bahrain",
  "Belarus",
  "Belgium",  
  "Bulgaria", 
  "Croatia", 
  "Cyprus",       
  "Czech Republic",              
  "Denmark",   
  "England", 
  "Finland",               
  "France",              
  "Germany",               
  "Greece", 
  "Greenland", 
  "Hungary",              
  "Iceland",
  "Ireland",                             
  "Italy", 
  "Latvia",        
  "Liechtenstein",            
  "Lithuania", 
  "Luxembourg", 
  "Netherland Antilles",          
  "Netherlands", 
  "Norway",    
  "Poland",             
  "Portugal", 
  "Romania",
  "Serbia",              
  "Slovakia",             
  "Slovenia",
  "Spain",            
  "Sweden",          
  "Switzerland", 
  "Turkey",     
  "Turks and Caicos",
  "Ukraine",
  "Bosnia", 
  "Estonia",                
  "Faroe Islands", 
  "Macedonia",             
  "Malta",     
  "Moldova",               
  "Monaco", 
  "San Marino"  
  )] <- "Europe"

## Cleaning data for winner over time plots
plotly_win_df = 
  boston_df %>% 
    mutate(
      official_time = as.POSIXct(official_time, format = "%H:%M:%OS"), 
      year = as.factor(year),
      pace = as.POSIXct(pace, format = "%H:%M:%OS"),
      place_overall = as.numeric(place_overall)
    )
```

<br>

# Winning Times by Year

```{r}
plotly_win_df %>% 
  filter(overall == 1) %>% 
  plot_ly(x = ~year, y = ~format(official_time, "%H:%M:%OS"), 
          mode = 'lines', type = 'scatter',
          name = 'Boston Winners',
          hoverinfo = "text",
          text = 
            ~paste0("Name: ", display_name,
                    "\n", "Year: ", year,
            "\n", "Time: ", format(official_time, "%H:%M:%OS"),
            "\n", "Pace: ", format(pace, "%H:%M:%OS"))) %>% 
  layout(
    title = "Boston Marathon Winners by Year",
         xaxis = list(title = "Year"),
         yaxis = list(title = "Official Time"))
```

This plot shows that over time, the Boston marathon winners are getting faster. There also seems to be some data that was recorded incorectly. Jim Knaub and Franz Nietlispach are reported as having an official time of 1:22:17 and 1:25:59 but the fastest marathon ever run is 1:59 (though not recorded [officially](https://www.nytimes.com/2019/10/12/sports/eliud-kipchoge-marathon-record.html)).

The fastest time ever run (excluding the incorrect sub two hour times) for the Boston Marathon was run by Martin E Duffy in 1975, with a time of 2:04:54. 

<br>

## Winning Times by Year by Gender

```{r warning = FALSE}
plotly_win_df %>% 
  filter(gender_result == 1) %>% 
  mutate(
    gender = na_if(gender, "U"),
    gender = as.factor(gender), 
    gender = recode(gender, "F" = "Female", "M" = "Male")) %>% 
  group_by(gender) %>% 
  plot_ly(x = ~year, y = ~format(official_time, "%H:%M:%OS"), 
          color = ~gender,
          mode = 'lines', type = 'scatter',
          hoverinfo = "text",
          text = 
            ~paste0("Name: ", display_name,
                    "\n", "Year: ", year,
            "\n", "Time: ", format(official_time, "%H:%M:%OS"),
            "\n", "Pace: ", format(pace, "%H:%M:%OS"))) %>% 
  layout(
    title = "Boston Marathon Winners by Year and Binary Gender",
         xaxis = list(title = "Year"),
         yaxis = list(title = "Official Time"))
```

We can see that male and females winners seem to be following the same general trend of decreasing marathon winning times over the year, but that trend is much more pronounced in males than among females. We can also see when the Boston Marathon allowed women to compete in 1972, and that the first ever female winner was Nina Kuscsik, with a time of 3:10:26. The data from 1993 and 1995 seems to have errors, just as the mens' times. This data set shows that Jean Driscoll was the female winner for both years with times of 1:34:50 and 1:40:42. The current world record  holder for the fastest female marathon ran is Brigid Kosgei who ran it in 2:14:04 in [2019](https://www.worldathletics.org/records/by-discipline/road-running/marathon/outdoor/women).

<br>

# How do the Boston Marathon winners compare to the record marathon time for the same year? 

```{r}

records_html = 
  read_html("https://www.topendsports.com/sport/athletics/record-marathon.htm")


record_marathon =
  records_html %>% 
  html_nodes("table") %>% 
  html_table(fill = T) %>% 
  lapply(., function(x) setNames(x, c("time", "date", "athlete", "country", "marathon"))) 

marathon_records = 
record_marathon %>% 
  as.data.frame() %>% 
  mutate(
    time = as.POSIXct(time, format = "%H:%M:%OS"),
  ) %>% 
  separate(date, into = c("month", "day", "year")) %>% 
    mutate(year = as.numeric(year)) %>% 
    select(-month, -day)


marathon_records %>% 
  plot_ly(
    x = ~year, y = ~format(time, "%H:%M:%OS"),
    mode = 'lines', type = 'scatter',
    name = "Marathon Record Times",
    hoverinfo = "text", text = ~paste0(
        "Name: ", athlete, 
        "\n", "Year: ", year,
        "\n", "Time: ", format(time, "%H:%M:%OS"),
        "\n", "Marathon: ", marathon)) %>% 
  layout(
    title = "Record Marathon Times by Year",
         xaxis = list(title = "Year"),
         yaxis = list(title = "Official Time")) 
  
```






