---
title: "Weather Analysis"
output: html_document
---

```{r include=FALSE, warning=FALSE}
## Packages
library(tidyverse)
library(dplyr)
library(viridis)
library(ggridges)
library(patchwork)
library(readxl)
library(leaflet)
library(plotly)
library(scales)
library(ggplot2)
library(httr)
library(rvest)

## Data Importing and Cleaning  
years_1 <- c(1900:2012, 2014)
years_2 <- c(2015:2019)
importing_data = function(x){
 
  if(str_detect(x, str_c(years_1, collapse = "|"))) {
  read_csv(x, na = c("NULL", "", "0"), col_types = "cicccciiiicc") 
  } 
  
  else if(str_detect(x, str_c(years_2, collapse = "|"))){
    read_csv(x, na = c("NULL", "", "0"), col_types = "cccicccccccccccccccccciiiiccc")
  }
}

boston_df <- 
  tibble(list.files("data", full.names = TRUE)) %>% 
  setNames("file_name") %>% 
  mutate(data = map(file_name, importing_data)) %>% 
  unnest(data) %>% 
  mutate(year = readr::parse_number(file_name),
        city = coalesce(city, residence),
        display_name = str_replace_all(display_name, "[^a-zA-Z0-9]", " ")) %>% 
  mutate(country_residence = replace(country_residence, country_residence == "AHO", "Netherland Antilles")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "ALB", "Albania")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "ALG", "Algeria")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "AND", "Andorra")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "ARG", "Argentina")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Argenti", "Argentina")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "AUS", "Australia")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Austral", "Australia")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "AUT", "Austria")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "BAH", "Bahamas")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "BAR", "Barbados")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Barbado", "Barbados")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "BDI", "Burundi")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "BLR", "Belarus")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "BEL", "Belgium")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "BER", "Bermuda")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "BRA", "Brazil")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "BRN", "Brunei")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "BUL", "Bulgaria")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "CAN", "Canada")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "CAY", "Cayman")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "CHI", "Chile")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "CHN", "China")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "COL", "Colombia")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Colombi", "Colombia")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "CRC", "Costa Rica")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Costa R", "Costa Rica")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "CRO", "Croatia")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "CYP", "Cyprus")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "CZE", "Czech Republic")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Czech R", "Czech Republic")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "DEN", "Denmark")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "DOM", "Dominican Republic")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Dominic", "Dominican Republic")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "ECU", "Ecuador")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "EGY", "Egypt")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "El Salv", "El Salvador")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "ESA", "El Salvador")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "ESP", "Spain")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "EST", "Estonia")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "ETH", "Ethiopia")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Ethiopi", "Ethiopia")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Faroe I", "Faroe Islands")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "FIN", "Finland")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "FLK", "Falkland Islands")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "FRA", "France")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "GBR", "England")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "GER", "Germany")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "GRE", "Greece")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "GRN", "Greenland")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "GUA", "Guatemala")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Guatema", "Guatemala")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "HKG", "Hong Kong")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Hong Ko", "Hong Kong")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "HON", "Honduras")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Hondura", "Honduras")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "HUN", "Hungary")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "INA", "Indonesia")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Indones", "Indonesia")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "IND", "India")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "IRL", "Ireland")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "ISL", "Iceland")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "ISR", "Israel")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "ITA", "Italy")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "JAM", "Jamaica")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "JPN", "Japan")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "JOR", "Jordan")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "KEN", "Kenya")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "KOR", "Korea")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Korea,", "Korea")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "KSA", "Saudi Arabia")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "KUW", "Kuwait")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "LAT", "Latvia")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "LIE", "Liechtenstein")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Liechte", "Liechtenstein")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Lithuan", "Lithuania")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "LTU", "Lithuania")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "LUX", "Luxembourg")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Luxembo", "Luxembourg")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Macao S", "Macao")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Macedon", "Macedonia")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Malaysi", "Malaysia")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "MAR", "Martinique")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Martini", "Martinique")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "MAS", "Malaysia")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "MEX", "Mexico")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "MGL", "Mongolia")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "MLT", "Malta")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "NCA", "Nicaragua")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "NED", "Netherlands")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Netherl", "Netherlands")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "New Zea", "New Zealand")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "NGR", "Nigeria")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "NOR", "Norway")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "NZL", "New Zealand")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "OMA", "Oman")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "PAK", "Pakistan")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Palesti", "Palestine")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "PAN", "Panama")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "PAR", "Paraguay")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Paragua", "Paraguay")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "PER", "Peru")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "PHI", "Philippines")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Philipp", "Philippines")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "POL", "Poland")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "POR", "Portugal")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Portuga", "Portugal")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Puerto", "Puerto Rico")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "QAT", "Qatar")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "ROU", "Romania")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Saudi A", "Saudi Arabia")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "SIN", "Singapore")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Singapo", "Singapore")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "SLO", "Slovenia")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Slovaki", "Slovakia")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Sloveni", "Slovenia")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "SMR", "San Marino")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "South A", "South Africa")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "SRB", "Serbia")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Sri Lan", "Sri Lanka")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "SUI", "Switzerland")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "SVK", "Slovakia")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "SWE", "Sweden")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Switzer", "Switzerland")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "TCA", "Turks and Caicos")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "THA", "Thailand")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Thailan", "Thailand")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "TPE", "Taipei")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "TRI", "Trinidad")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Trinida", "Trinidad")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "TUR", "Turkey")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "TWN", "Taiwan")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "UAE", "United Arab Emirates")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "UGA", "Uganda")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "UKR", "Ukraine")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "United", "United States")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "URU", "Uruguay")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "USA", "United States")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "VEN", "Venezuela")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Venezue", "Venezuela")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "VGB", "Virgin Islands")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "VIE", "Vietnam")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "ZIM", "Zimbabwe")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "RSA", "South Africa")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "RUS", "Russia")) %>%
  mutate(country_residence = replace(country_residence, country_residence == "Russian", "Russia")) %>%
  filter(!is.na(display_name)) %>% 
  select(-file_name, -residence, -first_name, -last_name) %>% 
  rename("country_citizenship" = "contry_citizenship" )

## Continent categorization
boston_df$continent[boston_df$country_residence %in% c(
  "Australia", 
  "New Zealand"
  )] <- "Australia"
boston_df$continent[boston_df$country_residence %in% c(
  "Algeria",
  "Burundi",
  "Egypt",
  "Ethiopia",
  "Kenya",  
  "Morocco", 
  "Nigeria,",  
  "South Africa", 
  "Uganda",
  "Zimbabwe"
  )] <- "Africa"
boston_df$continent[boston_df$country_residence %in% c(
 "China", 
  "Hong Kong", 
  "India",            
  "Indonesia",
  "Israel",  
  "Japan",               
  "Jordan", 
  "Korea",   
  "Malaysia",
  "Mongolia",              
  "Nepal",  
  "Oman", 
  "Pakistan",            
  "Palestine", 
  "Philippines", 
  "Qatar",   
  "Russia", 
  "Saudi Arabia",
  "Singapore",  
  "Sri Lanka",  
  "Taipei",               
  "Taiwan",            
  "Thailand",  
  "Vietnam", 
  "Brunei", 
  "Kuwait",                         
  "Macao",
  "United Arab Emirates"                                  
  )] <- "Asia"
boston_df$continent[boston_df$country_residence %in% c(
  "Canada",  
  "Mexico", 
  "Puerto Rico",
  "United States",
  "Bahamas",                        
  "Barbados",
  "Belize",              
  "Bermuda", 
  "Cayman",                
  "Dominican Republic",  
  "Jamaica", 
  "Martinique",
  "Virgin Islands"                                                           
  )] <- "North America"
boston_df$continent[boston_df$country_residence %in% c(
  "Argentina", 
  "Bolivia",                
  "Brazil", 
  "Chile", 
  "Colombia",           
  "Costa Rica",  
  "Ecuador",  
  "El Salvador", 
  "Guatemala",             
  "Honduras", 
  "Nicaragua",  
  "Panama",             
  "Paraguay", 
  "Peru",  
  "Uruguay",            
  "Venezuela",
  "Aruba",
  "Falkland Island",
  "Trinidad"
  )] <- "South America"
boston_df$continent[boston_df$country_residence %in% c(
  "Albania", 
  "Andorra",            
  "Armenia",
  "Austria",
  "Bahrain",
  "Belarus",
  "Belgium",  
  "Bulgaria", 
  "Croatia", 
  "Cyprus",       
  "Czech Republic",              
  "Denmark",   
  "England", 
  "Finland",               
  "France",              
  "Germany",               
  "Greece", 
  "Greenland", 
  "Hungary",              
  "Iceland",
  "Ireland",                             
  "Italy", 
  "Latvia",        
  "Liechtenstein",            
  "Lithuania", 
  "Luxembourg", 
  "Netherland Antilles",          
  "Netherlands", 
  "Norway",    
  "Poland",             
  "Portugal", 
  "Romania",
  "Serbia",              
  "Slovakia",             
  "Slovenia",
  "Spain",            
  "Sweden",          
  "Switzerland", 
  "Turkey",     
  "Turks and Caicos",
  "Ukraine",
  "Bosnia", 
  "Estonia",                
  "Faroe Islands", 
  "Macedonia",             
  "Malta",     
  "Moldova",               
  "Monaco", 
  "San Marino"  
  )] <- "Europe"

## Data cleaning for geo plots
boston_df2 = boston_df %>% 
  filter(year > 1999) %>% 
  filter(overall == 1) %>% 
  filter(gender == "M") %>% 
  select(year, city, state, overall, everything()) %>% 
  drop_na(city) %>% 
  separate(city, into = c("city", "state", "country"), sep = ",") %>% 
  select(-country, -state) %>% 
  writexl::write_xlsx("interactive_map_men.xlsx")
boston_df3 = boston_df  %>% 
  filter(year > 1999) %>% 
  filter(gender_result == 1) %>% 
  filter(gender == "F") %>% 
  select(year, city, state, overall, everything()) %>% 
  drop_na(city) %>% 
  writexl::write_xlsx("interactive_map_women.xlsx")

map_df = read_excel("data/latitude_longitude_winners.xlsx", sheet = 1) %>% 
  select(year, city, latitude, longitude, age, gender, official_time, display_name) %>% 
  rename(place = city) 
map_df2 = read_excel("data/latitude_longitude_winners.xlsx", sheet = 2) %>% 
  select(year, city, latitude, longitude, age, gender, official_time, display_name) %>% 
  rename(place = city) 

men_open = read_excel("data/geo_winners.xlsx", sheet = 4) %>% 
  janitor::clean_names() %>% 
  filter(!(year == 2013)) %>% 
  rename(place = country) %>% 
  separate(official_time, into = c("data", "official_time"), sep = " (?=[^ ]+$)") %>% 
  select(-data) 
lat_long = read_excel("data/lat_long.xlsx") %>% 
  select(-country) %>% 
  rename(place = name)
men_merge <- merge(men_open,lat_long,by="place") %>% 
  rename(display_name = name) %>% 
  mutate(gender = "M") 
men_merge2 = men_merge %>% 
  mutate(age = NA) %>% 
  filter(year > 1900) %>% 
  filter(year < 2000) 
men_total = rbind(men_merge2, map_df)
women_open = read_excel("data/geo_winners.xlsx", sheet = 3) %>% 
  janitor::clean_names() %>% 
  filter(!(year == 2013)) %>% 
  rename(place = country) %>% 
  separate(official_time, into = c("data", "official_time"), sep = " (?=[^ ]+$)") %>% 
  select(-data) 
women_merge = merge(women_open, lat_long, by="place") %>% 
  rename(display_name = name) %>% 
  mutate(gender = "F") 
  
women_merge2 = women_merge %>% 
  mutate(age = NA) %>% 
  filter(year > 1900) %>% 
  filter(year < 2000) 
woman_total = rbind(women_merge2, map_df2) %>% 
  drop_na(latitude)

women_wheelchair = read_excel("data/geo_winners.xlsx", sheet = "women_wheel_chair") %>% 
    janitor::clean_names() %>% 
    filter(!(year == 2013)) %>% 
    separate(official_time, into = c("data", "official_time"), sep = " (?=[^ ]+$)") %>% 
    select(-data) %>% 
    mutate(gender = "F")
men_wheelchair = read_excel("data/geo_winners.xlsx", sheet = "men_wheel_chair") %>% 
    janitor::clean_names() %>% 
    separate(official_time, into = c("data", "official_time"), sep = " (?=[^ ]+$)") %>% 
    select(-data) %>% 
    mutate(gender = "M")
wheelchair_merge = rbind(women_wheelchair, men_wheelchair) %>% 
  rename(place = country)
wheelchair_total = merge(wheelchair_merge, lat_long, by="place") 
winners_bind = rbind(women_merge, men_merge)

## Weather specific data cleaning
weather = read_excel("data/weather_conditions.xlsx") %>% 
  janitor::clean_names() %>% 
  filter(!(year == 2013)) %>% 
  separate(wind, into = c("wind_direction", "wind_speed"), sep = "\\s")  %>% 
  separate(wind_speed, into = c("wind_min", "wind_max"), sep = "-") %>% 
  mutate(wind_min = as.numeric(wind_min)) %>% 
  mutate(wind_max = as.numeric(wind_max)) %>% 
  mutate(wind_speed_average = (wind_min + wind_max)/2)
weather2 = weather %>% 
  select(boston_temp, year, wind_speed_average, sky)

winners_bind2 = winners_bind  %>% 
   filter(year > 1999) 
boston_df2 = boston_df %>% 
     filter(year > 1999) 
  
merge_weather2 <- merge(weather2,winners_bind2,by="year")
merge_weather_participants = merge(weather2,boston_df2,by="year") %>% 
  mutate(boston_temp = as.numeric(boston_temp)) %>% 
  mutate(seconds = as.numeric(seconds)) 
```

<br>

# Weather during Boston Marathons
```{r warning=FALSE}
plot4 = weather %>%
  mutate(text_label = 
           str_c("Wind Speed Average: ", wind_speed_average)) %>%
  plot_ly(x = ~year, y = ~boston_temp, text = ~text_label, color= ~ sky, size = ~wind_speed_average,
          type = "scatter", mode = "markers",  colors = "viridis", 
          sizes = c(50, 700), marker = list(opacity = 0.7)) %>%
  layout(
    title = 'Weather Over Time',
    xaxis = list(title = 'Year'),
    yaxis = list(title = 'Boston Temperature'), 
    legend = list(title=list(text='Sky Conditions')))
plot4
```
*Note datapoint size is related to average wind speed.

Over the past 20 years, the temperature in Boston during the marathons has generally stayed between 40-70 degrees and the sky conditions are usually clear. In 2012, there was a particularly hot marathon day with the temperature in the high 80's. Over the past few years, there have been relatively low wind speeds on average with exception to 2015. 

<br>

# Wind during Boston Marathons
```{r warning=FALSE}
ggplot(data = weather)+
  geom_segment(aes(x = year, xend = year, y = wind_min, yend = wind_max, colour = wind_direction), size = 5, alpha = 0.6) + 
  labs(
    title = "Wind Speed Range During Boston Marathons Over Time",
    x = "Year",
    y = "Wind Speed Range (mph)",
    color='Wind Direction')
```

In the early 2000's, the wind direction was generally toward the North/northeast and recently has been generally toward the west, northwest direction. There has been relatively low wind speed with little wind speed variability the past few years during the Boston marathon. Around 2010, there were relatively higher wind speeds. In 2007, there was a particularly windy marathon with speeds ranging from 20-30 mph.

<br>

# Comparison between Chamption Times and Weather 
```{r warning=FALSE}
plot5 = merge_weather2 %>%
  mutate(text_label = 
           str_c("Name: ", display_name, "\nYear: ", year, "\nResidence: ", place)) %>% 
 plot_ly(
    x = ~boston_temp, y = ~official_time,  text = ~text_label, color = ~sky, size = ~wind_speed_average,
          type = "scatter", mode = "markers",  colors = "viridis", 
          sizes = c(50, 700), marker = list(opacity = 0.7)) %>%
  layout(
    title = 'Champion Times and Weather',
    xaxis = list(title = "Boston's Temperature"),
    yaxis = list(title = 'Official Time'), 
    legend = list(title=list(text='Sky Conditions')))
plot5
```
*Note datapoint size is related to average wind speed.

The fastest marathon times in the past 20 years have occurred when the temperature is around 55 degrees. At temperatures 55 and below as well as 85 and above, there tends to be higher wind speeds. Although there are only a few data points above 70 degrees, one may hypothesize from the graph that as temperature increases, the fastest marathon time also increases. The fastest time was run was during medium wind speed, 55 degrees, and a clear sky. 

<br>

# Comparison between Participant Times and Weather 

```{r warning=FALSE}
plot6 = merge_weather_participants %>% 
   plot_ly(x = ~boston_temp, y = ~seconds, color = ~sky, type = "box") %>% 
  layout(
    title = 'Participant Times by Temperature',
    xaxis = list(title = "Boston's Temperature"),
    yaxis = list(title = 'Official Time (seconds)'), 
    legend = list(title=list(text='Sky Conditions')))
plot6
```

As temperature increases, the median marathon time among participants slightly increases. Furthermore, the lowest median marathon time occured around 55 degrees. According to a study called "Import of Environmental Parameters on Marathon Running Performance", the ideal tempature for professional runners is 37 degrees but for nonprofessional athletes is around 50 degrees. We also see that there are many outliers in the higher timepoints because there are likely a handful of very slow runners. 




